A graph-based data model for interactive evolutionary computation

### 2017-01-18 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Working on the Introduction a bit

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
This paper presents an alternative data model for the development of
Interactive Evolutionay Computation (IEC) algorithms. A graph %specify
                                %the elements of the graph - JJ
stored for efficiency in a graph database, is
proposed to model the overall system which consists of: a social network of
volunteers participating in the interactive evolutionary experiment, a
population of individuals in the evolutionary algorithm, 
the system\'s functionality and their interactions. This graph model can then be used
to tackle some of the issues associated with IEC such as, 
user fatigue, lack of volunteer participation, fitness estimation % what is this issue?
%Maybe the abstract should start with the issues we want to tackle as
%motivation. Motivation always goes first - JJ
and novelty. % These two last issues are not too clear, maybe they
             % could be explained in that motivation. 
The proposed graph database could be used to model IEC systems that 
follow a user-centered approach. % Graph is the model, graph database
                                % the implementation. Two different
                                % things. 
A case study is presented providing both 
conceptual and implementation details using the EvoSpace-Interactive
platform.
% It must prove that the issues have been addressed - JJ
\end{abstract}

@@

\section{Introduction}
% Maybe an introduction on why and where to use these types of
% algorithms? - jj
Interactive Evolutionary Algorithms (IEAs) are standard EAs whose
fitness evaluations are performed by persons within an interactive 
system.  Thus, the main loop of the EA needs the intervention of a
human for the quality assessment of each proposed solution. These
systems have an inherent drawback arising from the very nature of 
the algorithm, namely, the human fatigue caused by the interaction. % Can it be quantified? Maybe a reference to NodIO? - JJ
Some authors have already tried to address this basic problem
\cite{Frade:2010:EvoGAMES}. %Non sequitur. Is it something we are
                            %going to do here? - jj

% Maybe this claim should go to the beginning. THIS is why we want to
% use them - JJ
Nevertheless, IEAs have demonstrated their ability for effectively
producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992},
as well as other products in many other domains \cite{ie1}. 
Several IECs have been developed as web based systems \cite{picbreeder},
who depend on volunteers who visit the system to interactively help
with the search, using both anonymous and registered users. In particular
the EvoSpace-Interactive (ES-I) %citation
platform employs a collaborative technique,
where several registered users assign a quality assessment to a single
individual and then an aggregated fitness value is calculated. The fitness
assigned to each individual depends on the taste of each particular user, 
resulting in a many-to-many relationship between users and individuals. 
Many systems query this user-individual relation to extract relevant
knowledge about the process and the population, for instance showing the
most popular individual, or the the user with more participation \cite{picbreeder}.
In order to do this, individuals about individuals %what? Maybe data
                                %about individuals? - JJ
must be stored, even
if they are no longer in the active population. In this work a graph database
is proposed to model the interaction between users and individuals, with
the following advantages:

@@

Users are a central part of an IEC system,  not only because their
explicit evaluation is essential, but also because the overall context of interaction
affects the system as a whole as users interact not only with the graphical
interface, and each individual\'s phenotype, there is also the interaction
with other users, their previous evaluations and their own history of activities
and past experiences. Thus users in this proposal are dinamic entities 
interacting with an evolutionary computation and
have the fundamental purpose of evaluating individuals according to their preferences,
both explicitly or implicitly. Their degree of participation is affected by 
the context of use. With the porpouse of modeling the system, the contextual
factors are abstracted using a graph. This graph can then be used to 
enhance the system to capture the attention of users and possibly
increase their participation and mitigate the fatige.

@@

Currently there are some basic actions that
users can perform in the system, in order to interact with EvoSpace-I platform
users must login to the system through their facebook account. 
Once a user has logged-in to the system, he or she can do several actions.
Example of actions are the following:

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
This paper presents an alternative data model for the development of
Interactive Evolutionary Computation (IEC) algorithms. A graph %specify
                                %the elements of the graph - JJ
stored for efficiency in a graph database, is
proposed to model the overall system which consists of: a social network of
volunteers participating in the interactive evolutionary experiment, a
population of individuals in the evolutionary algorithm, 
the system\'s functionality and their interactions. This graph model can then be used
to tackle some of the issues associated with IEC such as, 
user fatigue, lack of volunteer participation, fitness estimation % what is this issue?
%Maybe the abstract should start with the issues we want to tackle as
%motivation. Motivation always goes first - JJ
and novelty. % These two last issues are not too clear, maybe they
             % could be explained in that motivation. 
The proposed graph database could be used to model IEC systems that 
follow a user-centered approach. % Graph is the model, graph database
                                % the implementation. Two different
                                % things. 
A case study is presented providing both 
conceptual and implementation details using the EvoSpace-Interactive
platform.
% It must prove that the issues have been addressed - JJ
\end{abstract}

@@

\section{Introduction}
% Maybe an introduction on why and where to use these types of
% algorithms? - jj
% Done -Mario
Interactive Evolutionary Algorithms (IEAs) are standard EAs whose
fitness evaluations are performed by persons within an interactive 
system.  Thus, the main loop of the EA needs the intervention of a
human for the quality assessment of each proposed solution.
IEAs have demonstrated their ability for effectively
producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992},
as well as other products in many other domains \cite{ie1}. 
Several IECs have been developed as web based systems \cite{picbreeder},
who depend on volunteers who visit the system to interactively help
with the search, using both anonymous and registered users.
These systems have several inherent drawbacks arising from the very nature of 
the algorithm, namely, the human fatigue caused by the interaction,
boredom or lack of interest when evaluating a large number of individuals,
many of which are not interesting or are very similar to each other.
IEAs share some of the issues commonly found in volunteer based systems
\cite{sarmenta2001volunteer,web:BOINC} such as volunteer\'s lack of accountability,
the need to build trust between participants and application providers. 
There is also the difficulty of establishing the amount of time and resources
a volunteer is willing to spend on the system, and how they decide if they
participate or not \cite{JJ:2016}.

 % Can it be quantified? Maybe a reference to NodIO? - JJ
Some authors have already tried to address this basic problem
\cite{Frade:2010:EvoGAMES}. %Non sequitur. Is it something we are
                            %going to do here? - jj

% Maybe this claim should go to the beginning. THIS is why we want to
% use them - JJ
 In particular
the EvoSpace-Interactive (ES-I) %citation
platform employs a collaborative technique,
where several registered users assign a quality assessment to a single
individual and then an aggregated fitness value is calculated. The fitness
assigned to each individual depends on the taste of each particular user, 
resulting in a many-to-many relationship between users and individuals. 
Many systems query this user-individual relation to extract relevant
knowledge about the process and the population, for instance showing the
most popular individual, or the the user with more participation \cite{picbreeder}.
In order to do this, individuals about individuals %what? Maybe data
                                %about individuals? - JJ
must be stored, even
if they are no longer in the active population. In this work a graph database
is proposed to model the interaction between users and individuals, with
the following advantages:

@@

Users are a central part of an IEC system,  not only because their
explicit evaluation is essential, but also because the overall context of interaction
affects the system as a whole as users interact not only with the graphical
interface, and each individual\'s phenotype, there is also the interaction
with other users, their previous evaluations and their own history of activities
and past experiences. Thus users in this proposal are dynamic entities 
interacting with an evolutionary computation and
have the fundamental purpose of evaluating individuals according to their preferences,
both explicitly or implicitly. Their degree of participation is affected by 
the context of use. With the purpose of modeling the system, the contextual
factors are abstracted using a graph. This graph can then be used to 
enhance the system to capture the attention of users and possibly
increase their participation and mitigate the fatigue.

@@

Currently there are some basic actions that
users can perform in the system, in order to interact with EvoSpace-I platform
users must login to the system through their Facebook account. 
Once a user has logged-in to the system, he or she can do several actions.
Example of actions are the following:
